---
title: "R Notebook"
output: html_notebook
---
```{r}
#Installation:
install.packages("tidyverse")
```


```{r}
#Necessary libraries:
library(tidyverse)

```


```{r}
#let's analyize the iris dataset:
head(iris) #show the first ten entries of the iris dataset

plot(iris$Sepal.Length,iris$Sepal.Width, xlab = "Sepal Length (cm)", ylab = "Sepal Width (cm)")

#using ggplot:
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width,color = Species)) +
 geom_point(aes(shape = Species)) + #changes the shape for each species
  scale_color_manual(values = c("purple",
                                "black",
                                "orange")) +             
  # geom_line() + #connects the dots
  geom_smooth() + #can add se=FALSE to remove the confidance margin
  # facet_wrap(~ Species)
facet_grid(rows = vars(Species))
  
#otheroptions - geom_boxplot, geom_col (coumn chart)

#now tht we seperated the data into species, the data eems correlated, which makes more sense. 

```

```{r}
#adding a grouping variable to iris:
iris2  <- iris
iris2$color <- rep(c("purple","red","black"), 50)
head(iris2)

ggplot(data = iris2, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point(size = 3) + 
  geom_smooth(size = 1, se = FALSE) + 
  facet_grid(rows = vars(Species), cols = vars(color))


```

```{r}
#diamonds dataset:
head(diamonds)

ggplot(diamonds, aes(carat, price)) + 
  geom_point(alpha = 0.05)+
  # scale_x_log10() +
    scale_y_continuous(trans = "log") + 
  scale_x_continuous(trans = "log")

#histogram: for numerical values n the x axis)
ggplot(diamonds, aes(price)) + 
  geom_histogram()

#bar chart - for catagorical values in the x axis:
ggplot(diamonds, aes(cut, fill = clarity)) + #fill adds the option of seperating the data in the bars by catagory
  geom_bar(position = "dodge") #seperates the different catagories into adjacent bars

#box plot - it draws a plot with a line at the data’s median, box borders at the 25% and 75% values, and lines reaching to the 5% and 95% values.

ggplot(diamonds, aes(cut, price)) + 
  geom_boxplot()


```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
